### 2024-03-05 types_2024/abstract.tex
stuff

% easychair.tex,v 3.5 2017/03/15

\documentclass{easychair}
%\documentclass[EPiC]{easychair}
%\documentclass[EPiCempty]{easychair}
%\documentclass[debug]{easychair}
%\documentclass[verbose]{easychair}
%\documentclass[notimes]{easychair}
%\documentclass[withtimes]{easychair}
%\documentclass[a4paper]{easychair}
%\documentclass[letterpaper]{easychair}

\usepackage{doc}
\usepackage{proof}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{newtx}

% use this if you have a long article and want to create an index
% \usepackage{makeidx}

% In order to save space or manage large tables or figures in a
% landcape-like text, you can use the rotating and pdflscape
% packages. Uncomment the desired from the below.
%
% \usepackage{rotating}
% \usepackage{pdflscape}

%% Front Matter
%%
% Regular title as in the article class.
%
\title{Type theory in type theory using single substitutions}

% Authors are joined by \and. Their affiliations are given by \inst, which indexes
% into the list defined using \institute
%
\author{
Ambrus Kaposi\inst{1}
\and
Szumi Xie\inst{1}
}

% Institutes for affiliations are also joined by \and,
\institute{
   Eötvös Loránd University,
   Budapest, Hungary, 
   \email{\{akaposi|szumi\}@inf.elte.hu}
}

%  \authorrunning{} has to be set for the shorter version of the authors' names;
% otherwise a warning will be rendered in the running heads. When processed by
% EasyChair, this command is mandatory: a document without \authorrunning
% will be rejected by EasyChair

\authorrunning{Kaposi, Xie}

% \titlerunning{} has to be set to either the main title or its shorter
% version for the running heads. When processed by
% EasyChair, this command is mandatory: a document without \titlerunning
% will be rejected by EasyChair
\titlerunning{Type theory in type theory using single substitutions}

\begin{document}

\maketitle

\newcommand{\Ra}{\Rightarrow}
\newcommand{\ra}{\rightarrow}
\newcommand{\Con}{\mathsf{Con}}
\newcommand{\Var}{\mathsf{Var}}
\newcommand{\var}{\mathsf{var}}
\newcommand{\Tm}{\mathsf{Tm}}
\newcommand{\Ty}{\mathsf{Ty}}
\newcommand{\Sub}{\mathsf{Sub}}
\newcommand{\U}{\mathsf{U}}
\newcommand{\El}{\mathsf{El}}
\renewcommand{\c}{\mathsf{c}}
\newcommand{\lam}{\mathsf{lam}}
\newcommand{\app}{\mathsf{app}}
\newcommand{\oldapp}{\mathbin{\cdot}}
\renewcommand{\tt}{\mathsf{tt}}
\newcommand{\Br}{\mathsf{Br}}
\newcommand{\aps}{\mathsf{aps}}
\newcommand{\ap}{\mathsf{ap}}
\newcommand{\mkBrPi}{\mathsf{mkBr}{\Pi}}
\newcommand{\Gel}{\mathsf{Gel}}
\newcommand{\gel}{\mathsf{gel}}
\newcommand{\ungel}{\mathsf{ungel}}
\renewcommand{\sp}{\hspace{1.5em}}
\newcommand{\blank}{\mathord{\hspace{1pt}\text{--}\hspace{1pt}}} %from the book
\newcommand{\Set}{\mathsf{Set}}
\newcommand{\ext}{\mathop{\triangleright}}
\newcommand{\R}{\mathsf{R}}
\newcommand{\p}{\mathsf{p}}
\renewcommand{\S}{\mathsf{S}}
\newcommand{\vz}{\mathsf{vz}}
\newcommand{\vs}{\mathsf{vs}}

In this abstract we define type theory in a minimalistic way: we aim
for an intrinsic, that is, well-typed and quotiented definition of the
syntax, that is as small as possible in terms of needed operations and
equations for the substitution calculus. We only introduce operations
that are \emph{forced} upon us. We aim for a language with $\Pi$
types, and it turns out that all the essential ingredients are forced
upon us when we want to introduce the usual rules for $\Pi$ types
(abstraction, application, $\beta$ and $\eta$).

We need variables in our language, so we introduce sorts of contexts,
types and variables.
\[
 \Con : \Set \hspace{6em} \Ty : \Con\ra\Set \hspace{6em} \Var : (\Gamma:\Con)\ra\Ty\,\Gamma\ra\Set  \\
\]
Contexts are either empty or extended with a type.
\[
\diamond : \Con \hspace{6em} \blank\ext\blank : (\Gamma:\Con)\ra\Ty\,\Gamma\ra\Con
\]
We define variables as well-typed De Bruijn indices, but to express
these we need to weaken types: e.g.\ the zero De Bruijn index $\vz$
has a weakened type. We introduce a new sort for substitutions $\Sub$,
am instantiation operation $\blank[\blank]$ on types, and a weakening
substitution $\p$. For now, $\Sub$ seems like an overkill because we
are only using $\blank[\p]$, but it will come handy soon.
\[
\Sub: \Con\ra\Con\ra\Set \hspace{3em} \blank[\blank] : \Ty\,\Gamma\ra\Sub\,\Delta\,\Gamma\ra\Ty\,\Delta \hspace{3em} \p : \Sub\,(\Gamma\ext A)\,\Gamma \hspace{3em}
\]
\[
\vz : \Var\,(\Gamma\ext A)\,(A[\p]) \hspace{3em} \vs : \Var\,\Gamma\,A\ra\Var\,(\Gamma\ext B)\,(A[\p])
\]
Now we introduce $\Pi$ types together with an equation on how
instantiation with $\p$ acts on them. This is tricky: as $\Pi$ binds a
new variable in its second argument, we need a new version of the
weakening substitution which leaves the last variable untouched. This
is why we introduce lifting of a substitution $\blank^{+}$, and now we
can state a general instantiation rule for $\Pi$ which works not only
for $\p$, but arbitrary substitutions.
\begin{alignat*}{10}
& \Pi:(A:\Ty\,\Gamma)\ra\Ty\,(\Gamma\ext A)\ra\Ty\,\Gamma \hspace{3em} && \blank^+ : (\gamma:\Sub\,\Delta\,\Gamma)\ra\Sub\,(\Delta\ext A[\gamma])\,(\Gamma\ext A) \\
& \Pi[] : (\Pi\,A\,B)[\gamma] = \Pi\,(A[\gamma])\,(B[\gamma^+])
\end{alignat*}
In addition to having variables, we need a sort of terms which
includes variables and lambda abstraction.
\[
\Tm : (\Gamma:\Con)\ra\Ty\,\Gamma\ra\Set \hspace{1.8em} \var : \Var\,\Gamma\,A\ra\Tm\,\Gamma\,A \hspace{1.8em} \lam : \Tm\,(\Gamma\ext A)\,B\ra\Tm\,\Gamma\,(\Pi\,A\,B)  \\
\]
To express application, we need single substitutions as well because
the argument of the function appears in the return type. In addition
to $\p$ and $\blank^+$, $\langle\blank\rangle$ is the third and last
operation for creating substitutions.
\[
\langle\blank\rangle : \Tm\,\Gamma\,A\ra\Sub\,\Gamma\,(\Gamma\ext A) \hspace{5em} \blank\cdot\blank: \Tm\,\Gamma\,(\Pi\,A\,B)\ra(a:\Tm\,\Gamma\,A)\ra\Tm\,\Gamma\,(B[\langle a\rangle])
\]
Now we would like to express the $\beta$ computation rule, but for
this we also need to be able to instantiate terms (in addition to
types).
\[
\blank[\blank] : \Tm\,\Gamma\,A\ra(\gamma:\Sub\,\Delta\,\Gamma)\ra\Tm\,\Delta\,(A[\gamma]) \hspace{3em} \Pi\beta : \lam\,t\cdot a = t[\langle a\rangle] 
\]
Now that we introduced instantiation of terms, we have to revisit all
of our terms and provide rules on how to instantiate them: first of
all, we need instantiation rules for $\lam$ and
$\blank\cdot\blank$. The rule $\lam[]$ is well-typed because of
$\Pi[]$, however ${\cdot}[]$ is not well-typed on its own and requires
a new equation $[\langle\rangle]$.
\[
\lam[] : (\lam\,t)[\gamma] = \lam\,(t[\gamma^+]) \hspace{1em} [\langle\rangle] : A[\langle a\rangle][\gamma] = A[\gamma^+][\langle a[\gamma]\rangle] \hspace{1em} {\cdot}[] : (t\cdot a)[\gamma] = (t[\gamma])\cdot(a[\gamma])
\]
Then we need instantiation rules for variables, we list these by case
distinction on the substitution: weakening of a variable increases the
index by one. When instantiating with lifted substitutions and single
substitutions, we have to do case distinction on the De Bruijn index
whether it is zero or successor. For the latter two cases, we need
type equations (named $[\p][^+]$ and $[\p][\langle\rangle]$) to
typeckeck the term equations.\vspace{-1.5em}
\begin{alignat*}{10}
  & && \var\,x[\p] = \var\,(\vs\,x) \\
  & [\p][^+] : A[\p][\gamma^+] = A[\gamma][\p] \hspace{4em}&& \var\,\vz[\gamma^+] = \var\,\vz \hspace{4em} && \var\,(\vs\,x)[\gamma^+] = \var\,x[\gamma][\p] \\
  & [\p][\langle\rangle] : A[\p][\langle a\rangle] = A && \var\,\vz[\langle a\rangle] = a &&  \var\,(\vs\,x)[\langle a\rangle] = \var\,x
\end{alignat*}
Finally, to typecheck the $\Pi\eta$ rule, we need our last equation on
types.
\[
[^+][\langle\rangle] : A[p^+][\langle\var\,\vz\rangle] = A \hspace{6em}\Pi\eta : t = \lam\,(t[\p]\cdot\var\,\vz)
\]
This concludes all the rules for type theory with $\Pi$: we summarise
as follows: there are three kinds of substitutions (single weakening,
single substitution, lifted substitution), we have instantiation rule
for each operation, there are 5 equations describing how instantiation
acts on variables, and 4 equations which describe general properties
of instantiation on types. The rest of the rules are specific to our
single type former $\Pi$: the only extra requirement is that each
operation should come equipped with an instantiation rule ($\Pi[]$,
$\lam[]$, ${\cdot}[]$). Perhaps surprisingly, this is enough to define
the syntax: there is no need for $\Con$ and $\Sub$ to a form a
category, no need for parallel substitutions, empty substitution,
parallel weakenings, or combinations of these.

We can add more type formers...

Show how to derive CwF.

\label{sect:bib}
\bibliographystyle{plainurl}
%\bibliographystyle{alpha}
%\bibliographystyle{unsrt}
%\bibliographystyle{abbrv}
\bibliography{b}

%------------------------------------------------------------------------------
\end{document}