### 2024-03-05 types_2024/abstract.tex
stuff

% easychair.tex,v 3.5 2017/03/15

\documentclass{easychair}
%\documentclass[EPiC]{easychair}
%\documentclass[EPiCempty]{easychair}
%\documentclass[debug]{easychair}
%\documentclass[verbose]{easychair}
%\documentclass[notimes]{easychair}
%\documentclass[withtimes]{easychair}
%\documentclass[a4paper]{easychair}
%\documentclass[letterpaper]{easychair}

\usepackage{doc}
\usepackage{proof}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{newtx}

% use this if you have a long article and want to create an index
% \usepackage{makeidx}

% In order to save space or manage large tables or figures in a
% landcape-like text, you can use the rotating and pdflscape
% packages. Uncomment the desired from the below.
%
% \usepackage{rotating}
% \usepackage{pdflscape}

%% Front Matter
%%
% Regular title as in the article class.
%
\title{Type theory in type theory using single substitutions}

% Authors are joined by \and. Their affiliations are given by \inst, which indexes
% into the list defined using \institute
%
\author{
Ambrus Kaposi
\and
Szumi Xie
}

% Institutes for affiliations are also joined by \and,
\institute{
   Eötvös Loránd University,
   Budapest, Hungary, 
   \email{\{akaposi|szumi\}@inf.elte.hu}
}

%  \authorrunning{} has to be set for the shorter version of the authors' names;
% otherwise a warning will be rendered in the running heads. When processed by
% EasyChair, this command is mandatory: a document without \authorrunning
% will be rejected by EasyChair

\authorrunning{Kaposi, Xie}

% \titlerunning{} has to be set to either the main title or its shorter
% version for the running heads. When processed by
% EasyChair, this command is mandatory: a document without \titlerunning
% will be rejected by EasyChair
\titlerunning{Type theory in type theory using single substitutions}

\begin{document}

\maketitle

\newcommand{\Ra}{\Rightarrow}
\newcommand{\ra}{\rightarrow}
\newcommand{\Con}{\mathsf{Con}}
\newcommand{\Var}{\mathsf{Var}}
\newcommand{\var}{\mathsf{var}}
\newcommand{\Tm}{\mathsf{Tm}}
\newcommand{\Ty}{\mathsf{Ty}}
\newcommand{\Sub}{\mathsf{Sub}}
\newcommand{\U}{\mathsf{U}}
\newcommand{\El}{\mathsf{El}}
\renewcommand{\c}{\mathsf{c}}
\newcommand{\lam}{\mathsf{lam}}
\newcommand{\app}{\mathsf{app}}
\newcommand{\oldapp}{\mathbin{\cdot}}
\renewcommand{\tt}{\mathsf{tt}}
\newcommand{\Br}{\mathsf{Br}}
\newcommand{\aps}{\mathsf{aps}}
\newcommand{\ap}{\mathsf{ap}}
\newcommand{\mkBrPi}{\mathsf{mkBr}{\Pi}}
\newcommand{\Gel}{\mathsf{Gel}}
\newcommand{\gel}{\mathsf{gel}}
\newcommand{\ungel}{\mathsf{ungel}}
\renewcommand{\sp}{\hspace{1.5em}}
\newcommand{\blank}{\mathord{\hspace{1pt}\text{--}\hspace{1pt}}} %from the book
\newcommand{\Set}{\mathsf{Set}}
\newcommand{\ext}{\mathop{\triangleright}}
\newcommand{\R}{\mathsf{R}}
\newcommand{\p}{\mathsf{p}}
\renewcommand{\S}{\mathsf{S}}
\newcommand{\vz}{\mathsf{vz}}
\newcommand{\vs}{\mathsf{vs}}

In this abstract we define type theory in a minimalistic way: we aim
for an intrinsic, that is, well-typed and quotiented definition of the
syntax, that is as small as possible in terms of number of operations
and equations for the substitution calculus. We only introduce
operations that are forced upon us. We aim for a language with $\Pi$
types, and it turns out that we have to add all the ingredients of the
substitution calculus while introducing the usual rules for $\Pi$
types (abstraction, application, $\beta$ and $\eta$). When adding more
type formers, we don't need to extend the substitution calculus
anymore.

We need variables in our language, so we introduce sorts of contexts,
types and variables.
\[
 \Con : \Set \hspace{6em} \Ty : \Con\ra\Set \hspace{6em} \Var : (\Gamma:\Con)\ra\Ty\,\Gamma\ra\Set  \\
\]
Contexts are either empty or are built from a context extended with a
type.
\[
\diamond : \Con \hspace{6em} \blank\ext\blank : (\Gamma:\Con)\ra\Ty\,\Gamma\ra\Con
\]
We define variables as well-typed De Bruijn indices, but to express
these we need to weaken types: e.g.\ the zero De Bruijn index $\vz$
has a weakened type. We introduce a new sort for substitutions $\Sub$,
an instantiation operation $\blank[\blank]$ on types, and a weakening
substitution $\p$. For now, $\Sub$ seems like an overkill because we
are only using $\blank[\p]$, but it will come handy soon.
\[
\Sub: \Con\ra\Con\ra\Set \hspace{3em} \blank[\blank] : \Ty\,\Gamma\ra\Sub\,\Delta\,\Gamma\ra\Ty\,\Delta \hspace{3em} \p : \Sub\,(\Gamma\ext A)\,\Gamma \hspace{3em}
\]
\[
\vz : \Var\,(\Gamma\ext A)\,(A[\p]) \hspace{3em} \vs : \Var\,\Gamma\,A\ra\Var\,(\Gamma\ext B)\,(A[\p])
\]
Now we introduce $\Pi$ types together with an equation on how
instantiation with $\p$ acts on them. This is tricky: as $\Pi$ binds a
new variable in its second argument, we need a new version of the
weakening substitution which leaves the last variable untouched. This
is why we introduce lifting of a substitution $\blank^{+}$, and now we
can state a general instantiation rule for $\Pi$ which works not only
for $\p$, but arbitrary substitutions (including lifted ones).
\begin{alignat*}{10}
& \Pi:(A:\Ty\,\Gamma)\ra\Ty\,(\Gamma\ext A)\ra\Ty\,\Gamma \hspace{3em} && \blank^+ : (\gamma:\Sub\,\Delta\,\Gamma)\ra\Sub\,(\Delta\ext A[\gamma])\,(\Gamma\ext A) \\
& \Pi[] : (\Pi\,A\,B)[\gamma] = \Pi\,(A[\gamma])\,(B[\gamma^+])
\end{alignat*}
In addition to having variables, we need a sort of terms which
includes variables and lambda abstraction.
\[
\Tm : (\Gamma:\Con)\ra\Ty\,\Gamma\ra\Set \hspace{1.8em} \var : \Var\,\Gamma\,A\ra\Tm\,\Gamma\,A \hspace{1.8em} \lam : \Tm\,(\Gamma\ext A)\,B\ra\Tm\,\Gamma\,(\Pi\,A\,B)  \\
\]
To express application, we need single substitutions as well because
the argument of the function appears in the return type. In addition
to $\p$ and $\blank^+$, $\langle\blank\rangle$ is the third and last
operation for creating substitutions.
\[
\langle\blank\rangle : \Tm\,\Gamma\,A\ra\Sub\,\Gamma\,(\Gamma\ext A) \hspace{5em} \blank\cdot\blank: \Tm\,\Gamma\,(\Pi\,A\,B)\ra(a:\Tm\,\Gamma\,A)\ra\Tm\,\Gamma\,(B[\langle a\rangle])
\]
Now we would like to express the $\beta$ computation rule, but for
this we also need to be able to instantiate terms (in addition to
types).
\[
\blank[\blank] : \Tm\,\Gamma\,A\ra(\gamma:\Sub\,\Delta\,\Gamma)\ra\Tm\,\Delta\,(A[\gamma]) \hspace{3em} \Pi\beta : \lam\,t\cdot a = t[\langle a\rangle] 
\]
Now that we have instantiation of terms, we need to revisit all
operations producing terms and provide rules on how to instantiate
them: first of all, we need instantiation rules for $\lam$ and
$\blank\cdot\blank$. The rule $\lam[]$ is well-typed because of
$\Pi[]$, however ${\cdot}[]$ is not well-typed on its own and requires
a new equation $[\langle\rangle]$.
\[
\lam[] : (\lam\,t)[\gamma] = \lam\,(t[\gamma^+]) \hspace{1em} [\langle\rangle] : A[\langle a\rangle][\gamma] = A[\gamma^+][\langle a[\gamma]\rangle] \hspace{1em} {\cdot}[] : (t\cdot a)[\gamma] = (t[\gamma])\cdot(a[\gamma])
\]
Then we need instantiation rules for variables, we list these for each
possible substitution: weakening of a variable increases the index by
one; when instantiating with lifted substitutions and single
substitutions, we have to do case distinction on the De Bruijn index
whether it is zero or successor. For the latter two cases, we need
type equations (named $[\p][^+]$ and $[\p][\langle\rangle]$) to
typeckeck the term equations.
\begin{alignat*}{10}
  & && \var\,x[\p] = \var\,(\vs\,x) \\
  & [\p][^+] : A[\p][\gamma^+] = A[\gamma][\p] \hspace{4em}&& \var\,\vz[\gamma^+] = \var\,\vz \hspace{4em} && \var\,(\vs\,x)[\gamma^+] = \var\,x[\gamma][\p] \\
  & [\p][\langle\rangle] : A[\p][\langle a\rangle] = A && \var\,\vz[\langle a\rangle] = a &&  \var\,(\vs\,x)[\langle a\rangle] = \var\,x
\end{alignat*}
Finally, to typecheck the $\Pi\eta$ rule, we need our last equation on
types.
\[
[\p^+][\langle\vz\rangle] : A[\p^+][\langle\var\,\vz\rangle] = A \hspace{6em}\Pi\eta : t = \lam\,(t[\p]\cdot\var\,\vz)
\]
This concludes all the rules for type theory with $\Pi$. We summarise
as follows: there are three kinds of substitutions (single weakening,
single substitution, lifted substitution), we have 5 equations
describing how instantiation acts on variables, and 4 equations which
describe general properties of instantiation on types. The rest of the
rules are specific to our single type former $\Pi$: the only extra
requirement is that each operation is equipped with an instantiation
rule ($\Pi[]$, $\lam[]$, ${\cdot}[]$). Perhaps surprisingly, this is
enough to define the syntax: there is no need for $\Con$ and $\Sub$ to
a form a category, no need for parallel substitutions, empty
substitution, parallel weakenings, or other combinations of these.

When adding new type formers, we only need the rules for the type
former, and an extra instantiation (naturality) rule for each
operation. For example, a Coquand-universe can be added by $\U :
\Ty\,\Gamma$, $\El:\Tm\,\Gamma\,\U\ra\Ty\,\Gamma$,
$\c:\Ty\,\Gamma\ra\Tm\,\Gamma\,\U$, $\U\beta : \El\,(\c\,A) = A$,
$\U\eta : \c\,(\El\,a) = a$, and three instantiation rules (note that
we need indexing to avoid inconsistency). In \cite{sogat}, we showed
that any second-order generalised algebraic theory (SOGAT) has a
single substitution presentation.

In the syntax (initial model, quotient inductive-inductive type) of
the above theory, all the rules of categories with families (CwF
\cite{DBLP:journals/corr/abs-1904-00827}) are admissible. That is, by
induction on the single substitution syntax, we can define parallel
substitutions (lists of terms) which are composable and form a
category; we can define instantiation by parallel substitutions for
types and terms, these have the usual universal property of
comprehension. The main ingredient for this construction is the notion
of $\alpha$-normal form: a kind of normal form which is still
quotiented by $\Pi\beta$, $\Pi\eta$, but does not include explicit
substitutions. If a type is in $\alpha$-normal form, we know whether
it is $\Pi$, $\U$, or $\El$ of a term. If a term is in $\alpha$-normal
form, we know whether it is a variable, a $\lam$, an application or a
code for a type (note that any function on an $\alpha$-normal has to
respect $\Pi\beta$, $\Pi\eta$, $\U\beta$, $\U\eta$). We prove
$\alpha$-normalisation (every term has a unique $\alpha$-normal form)
and then use induction on $\alpha$-normal forms to define parallel
instantiation and prove its properties.

We formalised the single substitution calculus with an infinite
hierarchy of types closed under $\Pi$ and $\U$ as a quotient
inductive-inductive type in Agda, proved $\alpha$-normalisation for
it, and derived all the rules for the parallel substitution calculus
(CwF equipped with $\Pi$ types and $\U$).

It is clear that the rules for the single substitution calculus are
all derivable from the CwF-rules. The other direction is however not
true: there are more models of the single substitution calculus than
the parallel substitution calculus. The situation is analogous to the
relationship of combinatory logic and lambda calculus: the former has
more models, but the syntaxes are isomorphic
\cite{DBLP:conf/fscd/AltenkirchKSV23}. Our single substitution
calculus is a minimalistic version of B-systems
\cite{AHRENS_EMMENEGGER_NORTH_RIJKE_2023} (C-systems are their
parallel substitution analogue): B-systems include telescopes and more
equations than our single substitution calculus, making them
equivalent to C-systems.

\label{sect:bib}
\bibliographystyle{plainurl}
%\bibliographystyle{alpha}
%\bibliographystyle{unsrt}
%\bibliographystyle{abbrv}
\bibliography{b}

%------------------------------------------------------------------------------
\end{document}