### 2025-04-07 fscd/response.txt
stuff

We thank the reviewers for their constructive criticism.

----------------------------------------------------------
Response to Reviewer 1
----------------------------------------------------------

> 2. In various places you speak of "equivalent" systems, but it is
> not always clear to me what you mean by that. E.g. line 93, in what
> sense B-systems are equivalent to CwF? Certainly it is not an
> equivalence of categories, that one exists between Bsys and Csys. If
> you are only referring to the syntax, then it would be good to
> specify so. This happens again in lines 149-150.

TODO

> 7. line 233. It is not clear to me what you mean by: "We instantiate
> the type with the Sub[...] weakening."

We will remove this unclear sentence.

> 13. line 568-569. The sentence starting "as most sorts..." is a bit
> obscure to me: in Section 2 you speak of a sort of types and one of
> terms, etc.

Sorry, we should have said that most sorts in the new model are given
by specific term-sorts in the old model. We will fix this.

----------------------------------------------------------
Response to Reviewer 2
----------------------------------------------------------

> - Problem 11: What is a contextual isomorphism? Why are `F_Con` and
> `F_Sub` not required to be invertible?

A contextual isomorphism is an isomorphism between the contextual
cores of the two models (that is, when context and substitutions are
inductively generated from types and terms). In a model of type
theory, the important structure is carried by types and terms, and
contexts/substitutions are only there as helpful infrastructure
(scaffolding). This is why the notion of contextual isomorphism makes
sense: it ignores the possible differences between
contexts/substitutions. We will add a definition and more explanation.

> As mentioned in the summary above, the claim that for type theory
> with Π-types, Σ-types and a universe hierarchy, the theory of CwFs
> is "weakly equivalent" to the SSC, is not well-defined and not fully
> substantiated.

We agree that this is problematic: as we mention in line 618--619,
types and terms are isomorphic after the other roundtrip, but because
of the limitations of the single substitution calculus, it is not
clear how to define contextual cores. We will acknowledge this more
directly and will only say (in the abstract and introduction) that a
CwF structure can be defined on an SSC with Σ,Π,U.

> I think the second half of §2.1 is a bit sloppy: - For Σ-types, it
> is suddenly mentioned that naturalities in one direction are
> derivable, but the argument used seems to apply to any type former
> with η.

Yes, this indeed works for any type former with η, if it is specified
as an isomorphism (Π is not specified as such, this is why we need
admissibility, see response to the next comment).

> - On ln 400, an admissible equation is suddenly used.

In the beginning of subsection 2.1, we say that this subsection will
motivate the need for admissible equations, and this is why we use
some equations defined later. This illustrates a limitation of the SSC
formulation: although the SSC theory is very concise, it is too weak
to even define the Π-isomorphism from the traditional version of
application in an arbitrary model.

> ## Minor comments
> 
> - Somewhat important in my opinion: a theorem environment labelled
> "Problem" gives the impression of an *open* problem.

We used Vladimir Voevodsky's convention from his paper "A C-system
defined by a universe in a category" where problem/construction refers
to a proof relevant theorem/proof pair. That is, when a construction
is a theorem, it means that it has at most one proof. We will clarify.

> - The word "instantiation" is used often but not really introduced;

A substitution is a list of terms. We use instantiation for the action
of a substitution. We will clarify.

> - ln 456: "It is important ...": [...] since αβη-equality is encoded
> as metatheoretic equality, any predicate will preserve the
> conversion rules.

It is true that any predicate preserves αβη-equality, but we cannot
define a map into a non-truncated version of α-normal forms, because
that would expose differences between βη-equal terms. This is the
reason for truncation.

> - ln 568-569: Do you mean "most sort inhabitants"?

See answer to reviewer 1.

> 1. A lot of the presented material is formalized in Agda. A
> recurring theme is admissibility vs. derivability.

We directly encode admissible equations as "provable in the syntax
(initial model)" and derivability as provable in any model. Similarly
for admissible operations/sorts. One can say that in Agda, pattern
matching corresponds to admissibility, and adding a definition to a
record corrsponds to derivability. We don't use pattern matching
because we simply postulate initial models and use their eliminator
instead.

> 2. To what extent can the central result in the paper (problem 6) be
> factored through the following result/conjecture: For any CwF, the
> category of parallel substitutions is exactly presented (†) by the
> graph of single-variable weakenings and substitutions, subject to
> the equations in def. 1 / def. 7 (expressed not as equations on
> type/term actions but as equations on morphisms in the free category
> over that graph).

We believe that the reviewer's conjecture does not hold: TODO

Ehrhard is exactly SSC+moving_equations_to_morphisms+ε+comp+id+⁺functoriality.

----------------------------------------------------------
Response to Reviewer 3
----------------------------------------------------------

We agree that we don't see a direct application of the SSC technique,
but we think that such a surprisingly concise and previously unknown
definition of type theory deserves a publication. Also, the general
phenomenon that for certain theories, the syntaxes are equivalent but
not the models, is interesting (like the relationship of lambda
calculus and SK-combinator calculus), and we work out a similar
example. This calculus is a kind of "minimisation" of type theory --
and it might be the first step in this direction: in the conclusions,
we speculate about its potential use as a coherent syntax, but this is
so far only fictional.